\documentclass{article}
\usepackage{hyperref}
%% \usepackage{epigraph}
\usepackage{titling}
\title{How the Web Functions}
\author{Dmitri Brengauz}

\begin{document}
\begin{titlingpage}
\maketitle
\begin{abstract}
  This essay gives a brief introduction to what goes on ``behind the
  scenes'' when a user requests a web page in a browser.
  Specifically, it answer the following questions:
  \begin{enumerate}
  \item \label{itm:landing} When we hit \texttt{https://www.techtonicgroup.com/} what happens? Don’t
    focus too much on architecture (Monolithic, SOA, Microservices,
    etc.). Try to focus more on how the web functions.
  \item From start to finish, how does that data reach you to be rendered
      in the browser?
      - \hyperlink{DNS}{DNS}/TCP/IP
 
  \item What code is rendered in the browser
  \item What is the server-side code's main function?
  \item How many instances of the client-side assets (HTML, CSS, JS,
    Images, etc.) are created?
  \item How many instances of the server side code are available at any given time?
  \item What is runtime?
  \item How many instances of the databases connected to the server
    application are created?
  \end{enumerate}
\end{abstract}
\end{titlingpage}

\section{Introduction}

With all the magic surrounding the World Wide Web, it's easy to lose
sight of what it is and was designed to be: a document retrieval
system.  Behind the scenes, the web browser, the Internet, and the
server should work seamlessly together to retrieve and compose that
document for viewing without the user having to worry about how any of
that is done.


\section{Down the Stack, Accross the Internet, Up the Stack}

%% \epigraph{The name of a resource indicates what we seek, an address indicates
%%  where it is, and a route tells us how to get there.}{\textit{John Shoch}}

To send and recieve data, the web browser itself forms the top layer
of the Internet Protocol Suite.  This is a stack of appplications,
organized together into four abstraction layers, that provides a uniform
interface for all of the programs on a computer to communicate over
the Internet.  Each layer depends on the layer below to fulfil it's
function, but has no need to know how that function is achieved.

In this example, the web browser forms the Application Layer.  It
requires a data connection to send and recieve messages from the
computer where information is located, the server.
\hypertarget{DNS}{The} browser uses another application level program,
the Domain Name Service, DNS, to translate the human-readable address
(like \$SITE) into a numerical address (IP address) that is used by
the rest of the stack to find a route to the server send data back and
forth across it.


The web browser turns to the layer below it on the stack, the
Transport Layer.  Here, the application has a choice: does it need a
continuous session and verification that each bit of data arrived in
the correct order, or can the remote computer just broadcast a more
efficient stream of data, with the occasional bits of data being lost?
A streaming video program would chose the latter option, and use the
UDP protocol in the Transport Layer.  However, the protocol that the
browser uses to talk to the server, HTTP/1.1, requires a stable
connection with error checking, and thus the web browser makes the
first choice, and sends its message using the Transmisson Control
Protocol, or TCP. 

The job of TCP is to establish that the machine at the other end is
ready to recieve and transmit data (establish a \textit{session})
divide the data sent to it into chunks called segments. TCP adds a
sequence number to each segment in a \textit{header}, which is sort of
like a data envelope. These sequence numbers make sure the data stays
in the correct order, no matter in which order the segments were
received by the remote machine. TCP is also in charge of requesting
that missing segments be re-transmitted, and transmitting segments
that have failed to arrive.  TCP itself doesn't know how to send and
receive segments from the network, however. For that, it relies on the
layer below it, the Internet Layer.

The Internet layer, using, IP or Internet Protocol, bundles TCP
segments into network packets. These have their own header, containing
the destination Internet address, the source (return) address, as well
as a checksum to check the payload---the TCP created segment--for corruption.  


IP knows the address of the machine on its own network
known as the gateway, which routes packets to the outside world, the
Internet.  It uses these IP addresses to instruct the final layer, the
Link layer, where to physically transmit the packets.


The Link layer is what communicates with the hardware. It allows the
transmission of IP network packets over copper Ethernet wires, or
electromagnetic waves though the air---WiFi signals. Once the packets
make it to the gateway, the Link layer there sends the packets back up
to the IP layer, which examines them, and using routing protocols
determines where they should be sent next.

Packets hop across networks like this until they reach their
destination. There, the IP layer sends them back up to the Transport
layer, which checks the data for completeness and puts it in order,
and finally delivers it to the application in charge of handling web
browser requests, the web server.

\section{Server Side}
The Application Level program that handles requests from web browsers
is called a web server.  The handling of requests Modern web servers
can handle tens of thousands of simultaneous connections, depending on
the memory and network bandwidth available, although sites that expect
high traffic use a device that distributes traffic across multiple web
servers called a load balancer.  The content that a web server will
send back to a browser can be separated into two types: static and
dynamic.

Static content can be delivered to the user's browser exactly as
stored on the server.  If it has to be generated or processed, then it
is called dynamic.  A web server can read static content directly from
the memory cache, or permanent storage if no cache exists.

Requests that require processing need to use a program that interprets
a set of instructions as to how this processing should be handled.
The Apache web server includes such programs internally as modules.
By contrast the Nginx server calls on an external application server
like the JavaScript runtime environment, Node.js, or Phusion Passenger
for Ruby on Rails.  Regardless of implementation, the code run on a
server to send back a web response to the client from the server is
called server-side scripting, or just ``server side.''

A major part of server-side scripting is interacting with a database
to retrieve or edit structured data. The term ``database'' refers to a
set of physical files, while an ``instance'' in an Oracle environment
refers to the combination of software and memory used to access that
data. In Microsoft SQL Server an instance is a specific installation of SQL Server.
Multiple SQL Server instances are used to manage server resources as
each instance is allocated a specific amount of memory and CPU
resources.

Each database instance can handle multiple web server connections,
with one connection typically being necessary per request, although
high-traffic databases use a technique to accommodate multiple
requests per connection known as pooling.











\end{document}
